---
title: "STAA-566: Data Vis"
subtitle: "Assignment 4: tables"
author: "Penny Grant"
date: "`r Sys.Date()`"
header-includes:
  - \setmainfont{Noto Sans}
  - \setmonofont{Noto Sans Mono}
  - \usepackage{physics}
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r, message = FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(reactable)
```


# Import Data

For this assignment, I thought I would revisit the dataset I used for our last assignment. However, instead of using just the happiness score, I wanted 

This data set was published by the World Happiness Report and downloaded from [kaggle](https://www.kaggle.com/datasets/mathurinache/world-happiness-report). It uses information from the Gallup World Survey and bases scores off several different factors and the degree to which they contribute to the country's happiness.

Description for columns gathered from:

> Helliwell, J. F., Layard, R., Sachs, J. D., De Neve, J.-E., Aknin, L. B., & Wang, S. (Eds.). (2022). World Happiness Report 2022. New York: Sustainable Development Solutions Network. https://happiness-report.s3.amazonaws.com/2022/WHR+22.pdf, page 21.

```{r,warning=FALSE, message = FALSE}
happiness <- read_csv("2022happinessReport.csv", 
     col_types = cols(`Dystopia (1.83) + residual` = col_number(), 
         `Explained by: GDP per capita` = col_number(), 
         `Explained by: Social support` = col_number(), 
         `Explained by: Healthy life expectancy` = col_number(), 
         `Explained by: Freedom to make life choices` = col_number(), 
         `Explained by: Generosity` = col_number(), 
         `Explained by: Perceptions of corruption` = col_number()))
```

Many of the predictors were calculated as the result of the answers to a poll question. The related question is listed below.

* **GDP per Capita**: natural log of GDP
* **Social Support**: *If you were in trouble, do you have relatives or friends you can count on to help you whenever you need them, or not?*
* **Healthy Life Expectancy**: based on data from the World Health Organization Global Health Observatory
* **Freedom to make life choices**: *Are you satisfied or dissatisfied with your freedom to choose what you do with your life?*
* **Generosity**: regression of answers to *Have you donated money to a charity in the past month?* on GDP
* **Perception of Corruption**: average of answers to two questions:
  1. *Is corruption widespread throughout the government in this country or not?*
  2. *Is corruption widespread within businesses in this country or not?*


# Make a Table


```{r}
tbl_happy <- happiness %>% 
  # select columns for table
  select(RANK, Country, `Happiness score`,`Explained by: GDP per capita`,
         `Explained by: Social support`,
         `Explained by: Healthy life expectancy`,
         `Explained by: Freedom to make life choices`, 
         `Explained by: Generosity`,
         `Explained by: Perceptions of corruption`) %>%
  
  reactable(
    # make default settings for all columns
    defaultColDef = colDef(
      headerStyle = list(fontWeight = 750),
      align = "center",
      format = colFormat(separators = TRUE)
    ),
    
    # customize column names
    columns = list(
    RANK = colDef(name = "Ranking", align = "center"),
    `Explained by: GDP per capita`= colDef(name="GDP per capita"),
    `Explained by: Social support`= colDef(name = "Social Support"),
    `Explained by: Healthy life expectancy`= 
        colDef(name = "Healthy Life Expectancy"),
    `Explained by: Freedom to make life choices` = 
        colDef(name = "Freedom to make life choices"),
    `Explained by: Generosity`= colDef(name = "Generosity"),
    `Explained by: Perceptions of corruption` =
        colDef(name = "Perceptions of Corruption")
  ),
  columnGroups = list(
    colGroup(name = "Explained by",columns=c("Explained by: GDP per capita",
         "Explained by: Social support",
         "Explained by: Healthy life expectancy",
         "Explained by: Freedom to make life choices", 
         "Explained by: Generosity",
         "Explained by: Perceptions of corruption"))),
  searchable = TRUE,
  showPageSizeOptions = TRUE,
  striped = TRUE,
  style = list(fontSize = "1rem"),
  defaultSorted = "RANK")
```



```{r}
tbl_happy
```










